Perturbing SU($n$)-symmetric interactions

### 2020-02-05 collective_sun/sun_perturbation.tex
fixed minor error with diagram prefactors
changed the definition of an unlabeled diagram

Calculating the second-order effective Hamiltonian $H_1^{(2)}$ induced
on the permutationally symmetric manifold $\M_0$ by the perturbation
$\V_1$ in \eqref{eq:perturbations} requires us to simplify a product
of the form
\begin{align}
  \P_0 \sp{\sum_{p,q} v_p w_q X_p Y_q} \P_0
  \EQFS \diagram{single_body_2} X_1 Y_2
  + \diagram{single_body_1} X_1 Y_1,
  \label{eq:PXYP_start}
\end{align}
where $p,q$ index individual spins; $X,Y$ are single-spin operators;
and the coefficients $v_p,w_q$ satisfy \eqref{eq:cond_1} with
eigenvalue $\Delta\ne0$, which implies that they are mean-zero.
Defining $\v m \equiv \sum_p m_p \ket{p}$ for $m\in\set{v,w}$, we can
simplify
\begin{align}
  \diagram{single_body_1}
  \equiv \sum_p v_p w_p
  = \v v\c\v w,
\end{align}
and
\begin{align}
  \diagram{single_body_2}
  = \diagram{single_body_2_o} - \diagram{single_body_2_x}
  = - \diagram{single_body_1}
  = - \v v \c\v w,
\end{align}
so
\begin{align}
  \sum_{p,q} v_p w_q X_p Y_q
  \EQFS - \v v\c\v w \p{X_1 Y_2 - X_1 Y_1},
\end{align}
In order to write this result in terms of collective operators
$\col{Z} \equiv \sum_p Z_p$, we expand
\begin{align}
  \col{X Y} \EQFS N X_1 Y_1,
  &&
  \col{X}\,\col{Y}
  = \sum_{p=q} X_p Y_q + \sum_{p\ne q} X_p Y_q
  \EQFS \col{XY} + N\p{N-1} X_1 Y_2,
\end{align}
which implies that
\begin{align}
  \sum_{p,q} v_p w_q X_p Y_q
  \EQFS - \f{\v v \c\v w}{N\p{N-1}}
  \p{\col{X}\,\col{Y} - N \col{XY}}.
\end{align}

@@

Here we simplify a product of three two-body operators of a particular
form projected onto the permutationally symmetric manifold:
\begin{align}
  \X \equiv \P_0 \sp{\sum_{k<\ell} \sum_{p<q} \sum_{r<s}
    u_{k\ell} v_{pq} w_{rs} X_k X_\ell X_p X_q X_r X_s} \P_0,
  &&
  X^2 = 1,
\end{align}
where $k,\ell,p,q,r,s$ index individual spins; $X$ is a single-spin
operator; and $u_{k\ell},v_{pq},w_{rs}$ are arbitrary scalar
coefficients with $m_{pq}=m_{qp}$ and $m_{pp}=0$ for each of
$m\in\set{u,v,w}$.  Collecting terms according to the number of $X$
operators that remain after accounting for $X^2=1$, we find that
\begin{align}
  \X \EQFS A_6 X^{\otimes 6} + A_4 X^{\otimes 4} + A_2 X^{\otimes 2},
  \label{eq:triple_multi}
\end{align}
with
\begin{align}
  A_6 \equiv \diagram{triple_6},
  &&
  A_4 \equiv \diagram{triple_4_1} + \diagram{triple_4_0},
  \label{eq:triple_64}
\end{align}
\begin{align}
  A_2 \equiv \diagram{triple_2_0} + \diagram{triple_2_1}
  + \diagram{triple_2_2} + \diagram{triple_2_3},
  \label{eq:triple_2}
\end{align}
where a diagram with unlabeled circles implicitly contains a sum over
all label assignments that result in a distinct diagram, e.g.
\begin{align}
  \diagram{triple_6} \equiv \diagram{triple_6_uvw},
  \label{eq:triple_uvw}
\end{align}
\begin{align}
  \diagram{triple_2_3} \equiv \diagram{triple_2_3_uvw}
  + \diagram{triple_2_3_vwu} + \diagram{triple_2_3_wuv}.
  \label{eq:triple_uvw_3}
\end{align}
The first diagram in each of $A_6,A_4,A_2$ as written in
\eqref{eq:triple_64}, \eqref{eq:triple_2} thus has only one assignment
of labels, as in \eqref{eq:triple_uvw}, while all other unlabeled
diagrams have three assignments, as in \eqref{eq:triple_uvw_3}.  We
can write \eqref{eq:triple_multi} in terms of collective operators as
\begin{align}
  \X \EQFS
  \tilde A_6 \col{X}^6 + \tilde A_4 \col{X}^4 + \tilde A_2 \col{X}^2,
  \label{eq:triple_col}
\end{align}
where we have neglected scalar terms without physical consequence, and
\begin{align}
  \tilde A_6 \equiv \f{A_6}{N!_6},
  &&
  \tilde A_4 \equiv \f{A_4}{N!_4} - 5\p{3N-8} \f{A_6}{N!_6},
\end{align}
\begin{align}
  \tilde A_2 \equiv \f{A_2}{N!_2} - 2\p{3N-4} \f{A_4}{N!_4}
  + \p{45N^2-210N+184} \f{A_6}{N!_6}.
\end{align}

>>>>

Calculating the second-order effective Hamiltonian $H_1^{(2)}$ induced
on the permutationally symmetric manifold $\M_0$ by the perturbation
$\V_1$ in \eqref{eq:perturbations} requires us to simplify a product
of the form
\begin{align}
  \P_0 \sp{\sum_{p,q} v_p w_q X_p Y_q} \P_0
  \EQFS \diagram{single_body_2} X_1 Y_2
  + \diagram{single_body_1} X_1 Y_1,
  \label{eq:PXYP_start}
\end{align}
where $p,q$ index individual spins; $X,Y$ are single-spin operators;
and the coefficients $v_p,w_q$ satisfy \eqref{eq:cond_1} with
eigenvalue $\Delta\ne0$, which implies that they are mean-zero.
Defining $\v m \equiv \sum_p m_p \ket{p}$ for $m\in\set{v,w}$, we can
simplify
\begin{align}
  \diagram{single_body_1}
  \equiv \sum_p v_p w_p
  = \v v\c\v w,
\end{align}
and
\begin{align}
  \diagram{single_body_2}
  = \diagram{single_body_2_o} - \diagram{single_body_2_x}
  = \diagram{single_body_2_oo} - \diagram{single_body_1}
  = - \v v \c\v w,
\end{align}
so
\begin{align}
  \sum_{p,q} v_p w_q X_p Y_q
  \EQFS - \v v\c\v w \p{X_1 Y_2 - X_1 Y_1},
\end{align}
In order to write this result in terms of collective operators
$\col{Z} \equiv \sum_p Z_p$, we expand
\begin{align}
  \col{X Y} \EQFS N X_1 Y_1,
  &&
  \col{X}\,\col{Y}
  = \sum_{p=q} X_p Y_q + \sum_{p\ne q} X_p Y_q
  \EQFS \col{XY} + N\p{N-1} X_1 Y_2,
\end{align}
which implies that
\begin{align}
  \sum_{p,q} v_p w_q X_p Y_q
  \EQFS - \f{\v v \c\v w}{N\p{N-1}}
  \p{\col{X}\,\col{Y} - N \col{XY}}.
\end{align}

@@

Here we simplify a product of three two-body operators of a particular
form projected onto the permutationally symmetric manifold:
\begin{align}
  \X \equiv \P_0 \sp{\sum_{k<\ell} \sum_{p<q} \sum_{r<s}
    u_{k\ell} v_{pq} w_{rs} X_k X_\ell X_p X_q X_r X_s} \P_0,
  &&
  X^2 = 1,
\end{align}
where $k,\ell,p,q,r,s$ index individual spins; $X$ is a single-spin
operator; and $u_{k\ell},v_{pq},w_{rs}$ are arbitrary scalar
coefficients with $m_{pq}=m_{qp}$ and $m_{pp}=0$ for each of
$m\in\set{u,v,w}$.  Collecting terms according to the number of $X$
operators that remain after accounting for $X^2=1$, we find that
\begin{align}
  \X \EQFS A_6 X^{\otimes 6} + A_4 X^{\otimes 4} + A_2 X^{\otimes 2},
  \label{eq:triple_multi}
\end{align}
with
\begin{align}
  A_6 \equiv \f18 \diagram{triple_6},
  &&
  A_4 \equiv \diagram{triple_4_1} + \f32 \diagram{triple_4_0},
  \label{eq:triple_64}
\end{align}
\begin{align}
  A_2 \equiv \f12 \diagram{triple_2_0} + 3 \diagram{triple_2_1}
  + \f34 \diagram{triple_2_2} + 3 \diagram{triple_2_3},
  \label{eq:triple_2}
\end{align}
where a diagram with unlabeled circles denotes an average over all
label assignments that result in a distinct diagram, e.g.
\begin{align}
  \diagram{triple_6} \equiv \diagram{triple_6_uvw},
  \label{eq:triple_uvw}
\end{align}
\begin{align}
  \diagram{triple_2_3}
  \equiv \f13 \sp{\diagram{triple_2_3_uvw}
  + \diagram{triple_2_3_vwu} + \diagram{triple_2_3_wuv}}.
  \label{eq:triple_uvw_3}
\end{align}
The first diagram in each of $A_6,A_4,A_2$ as written in
\eqref{eq:triple_64}, \eqref{eq:triple_2} thus has only one assignment
of labels, as in \eqref{eq:triple_uvw}, while all other unlabeled
diagrams have three assignments, as in \eqref{eq:triple_uvw_3}.  We
can write \eqref{eq:triple_multi} in terms of collective operators as
\begin{align}
  \X \EQFS
  \tilde A_6 \col{X}^6 + \tilde A_4 \col{X}^4 + \tilde A_2 \col{X}^2,
  \label{eq:triple_col}
\end{align}
where we have neglected scalar terms without physical consequence, and
\begin{align}
  \tilde A_6 \equiv \f{A_6}{N!_6},
  &&
  \tilde A_4 \equiv \f{A_4}{N!_4} - 5\p{3N-8} \f{A_6}{N!_6},
\end{align}
\begin{align}
  \tilde A_2 \equiv \f{A_2}{N!_2} - 2\p{3N-4} \f{A_4}{N!_4}
  + \p{45N^2-210N+184} \f{A_6}{N!_6}.
\end{align}
